Community engagement and subgroup meta-knowledge: Some factors in the soul of a community

### 2015-08-05 PaperFinalDraft.Rnw
reformatting dplyr blocks to close #9, found one more ggplot2 block

<<missingdatatable, results="asis">>=
count_na <- function(x) sum(is.na(x))
missing10 <- count_na(sotc10$recommend)
missing09 <- count_na(sotc09$recommend)
missing08 <- group_by(sotc08, recommend == "(Refused)") %>% summarise(n = n()) 
missing <- c(missing08$n[2], missing09, missing10)
missingtable <- as.matrix(data.frame(Year=c(2008, 2009, 2010), Missing=missing, Total=c(dim(sotc08)[1], dim(sotc09)[1], dim(sotc10)[1]), Percent=missing/c(dim(sotc08)[1], dim(sotc09)[1], dim(sotc10)[1])*100))
rownames(missingtable) <- NULL
print(xtable(x=missingtable,caption="Percent of missing data from surveys. The 2010 survey has almost 25 percent missing data. The total number of entries in 2010 is also much larger, suggesting that the 2008 and 2009 datasets used a different criteria for inclusion of entries.", display=c("d", "d", "d","d", "f"), label="missingtable"),include.rownames=FALSE)
@

@@

<<minorities>>=
races10 <- filter(sotc10, race != " " & race != "4" & race != "Don't Know") %>% group_by(race) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2010)
races10$race <- factor(races10$race)
levels(races10$race) <- c(levels(races10$race)[1:5], "Some other race", levels(races10$race)[7])

races09 <- filter(sotc09, race != " " & race != "4" & race != "Don't know") %>% group_by(race) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2009)
races09$race <- factor(races09$race)

races08 <- filter(sotc08, race != " " & race != "4" & race != "(DK)" & race != "None") %>% group_by(race) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2008)
races08$race <- factor(races08$race)
levels(races08$race) <- c("Hispanic", "More than one", "Refused", "American Indian or Alaskan", "Asian", "Black or African-American", "Native Hawaiian or other Pacific Islander", "Some other race", "White")

@@

<<minoritiesOverYearsE>>=
minoritiesYN08 <- filter(sotc08, race != " " & race != "4" & race != "(DK)" & race != "None" & race != "(Refused)") %>% filter(minorities != "(DK)" & minorities != "(Refused)") %>% group_by(citystate, race != "White", minorities) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2008)
names(minoritiesYN08) <- n1

minoritiesYN09 <- filter(sotc09, race != " " & race != "4" & race != "Don't know" & race != "Refused") %>% filter(is.na(minorities)==FALSE) %>% group_by(citystate, race != "White", minorities) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2009)
names(minoritiesYN09) <- n1

minoritiesYN10 <- filter(sotc10, race != " " & race != "4" & race != "Don't Know" & race != "Refused") %>% filter(is.na(minorities)==FALSE) %>% group_by(citystate, race != "White", minorities) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2010)
names(minoritiesYN10) <- n1

@@

<<samplesize, out.width="0.99\\linewidth",fig.width = 8.5,fig.height=4.5, fig.cap="Sample sizes for plots about meta-knowledge regarding the community as a place for minorities.">>=
whiteYN1 <- tbl_df(whiteYN) %>% group_by(year, InGroup) %>% summarise(howmany=sum(n)) %>% mutate(frac=prop.table(howmany))
whiteplot <- ggplot(aes(x=year, y=howmany), data=whiteYN1)+geom_bar(aes(fill=InGroup), position="dodge", stat="identity")+xlab("Year")+ylab("Sample size")+guides(fill=guide_legend(title=NULL))+guides(fill=guide_legend(title=NULL))+scale_fill_manual(values=colors2, labels=c("White", "Non-white"))
whiteplot
@

@@

<<OverallSeniors>>=
seniorGroup10 <- filter(sotc10, is.na(seniors) == FALSE) %>% group_by(citystate, age >= 62, seniors) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2010)

@@

seniorGroup09 <- filter(sotc09, is.na(age) == FALSE) %>% group_by(citystate, age >= 62, seniors) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2009)
names(seniorGroup09) <- n1

seniorGroup08 <- group_by(sotc08, citystate, age >=62, seniors) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2008)
names(seniorGroup08) <- n1

@@

<<kidsbyYear>>=
kidGroup10 <- filter(sotc10, children == "Yes" | children == "No", is.na(families_kids) == FALSE) %>% group_by(citystate, children, families_kids) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2010)
names(kidGroup10) <- n1

kidGroup09 <- filter(sotc09, children == "Yes" | children == "No", is.na(families_kids) == FALSE) %>% group_by(citystate, children, families_kids) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2009)
names(kidGroup09) <- n1

@@

kidGroup08 <- filter(sotc08, children == "Yes" | children == "No", is.na(families_kids) == FALSE & families_kids != "(DK)" & families_kids != "(Refused)") %>% group_by(citystate, children, families_kids) %>% summarise(n = n()) %>% mutate(frac=prop.table(n), year=2008)
names(kidGroup08) <- n1

@@

<<kidsSS,out.width="0.99\\linewidth",fig.width = 8.5,fig.height=4.5, fig.cap="Sample sizes for plots about meta-knowledge regarding the community as a place for families with young children.">>=
kidSS <- group_by(kidYears, year, InGroup) %>% summarise(number=sum(num)) %>% mutate(percentage=100*number/sum(number))
levels(kidSS$InGroup) <- c("Don't know", "Refused", "Childless households", "Families with children")
kidsplot <- ggplot(aes(x=year, y=number), data=kidSS)+geom_bar(aes(fill=InGroup), position="dodge", stat="identity")+xlab("Year")+ylab("Sample size")+guides(fill=guide_legend(title=NULL))+scale_fill_manual(values=colors2)
kidsplot 
@

>>>>

<<missingdatatable, results="asis">>=
count_na <- function(x) sum(is.na(x))
missing10 <- count_na(sotc10$recommend)
missing09 <- count_na(sotc09$recommend)
missing08 <- sotc08 %>%
  group_by(recommend == "(Refused)") %>% 
  summarise(n = n()) 
missing <- c(missing08$n[2], missing09, missing10)
missingtable <- as.matrix(data.frame(Year=c(2008, 2009, 2010), Missing=missing, Total=c(dim(sotc08)[1], dim(sotc09)[1], dim(sotc10)[1]), Percent=missing/c(dim(sotc08)[1], dim(sotc09)[1], dim(sotc10)[1])*100))
rownames(missingtable) <- NULL
print(xtable(x=missingtable,caption="Percent of missing data from surveys. The 2010 survey has almost 25 percent missing data. The total number of entries in 2010 is also much larger, suggesting that the 2008 and 2009 datasets used a different criteria for inclusion of entries.", display=c("d", "d", "d","d", "f"), label="missingtable"),include.rownames=FALSE)
@

@@

<<minorities>>=
races10 <- sotc10 %>% 
  filter(race != " " & race != "4" & race != "Don't Know") %>% 
  group_by(race) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)
races10$race <- factor(races10$race)
levels(races10$race) <- c(levels(races10$race)[1:5], "Some other race", levels(races10$race)[7])

races09 <- sotc09 %>% 
  filter(race != " " & race != "4" & race != "Don't know") %>% 
  group_by(race) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
races09$race <- factor(races09$race)

races08 <- sotc08 %>% 
  filter(race != " " & race != "4" & race != "(DK)" & race != "None") %>% 
  group_by(race) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2008)
races08$race <- factor(races08$race)
levels(races08$race) <- c("Hispanic", "More than one", "Refused", "American Indian or Alaskan", "Asian", "Black or African-American", "Native Hawaiian or other Pacific Islander", "Some other race", "White")

@@

<<minoritiesOverYearsE>>=
minoritiesYN08 <- sotc08 %>% 
  filter(race != " " & race != "4" & race != "(DK)" & race != "None" & race != "(Refused)") %>% 
  filter(minorities != "(DK)" & minorities != "(Refused)") %>% 
  group_by(citystate, race != "White", minorities) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2008)
names(minoritiesYN08) <- n1

minoritiesYN09 <- sotc09 %>% 
  filter(race != " " & race != "4" & race != "Don't know" & race != "Refused") %>% 
  filter(is.na(minorities)==FALSE) %>% group_by(citystate, race != "White", minorities) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(minoritiesYN09) <- n1

minoritiesYN10 <- sotc10 %>% 
  filter(race != " " & race != "4" & race != "Don't Know" & race != "Refused") %>% 
  filter(is.na(minorities)==FALSE) %>% 
  group_by(citystate, race != "White", minorities) %>% summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)
names(minoritiesYN10) <- n1

@@

<<samplesize, out.width="0.99\\linewidth",fig.width = 8.5,fig.height=4.5, fig.cap="Sample sizes for plots about meta-knowledge regarding the community as a place for minorities.">>=
whiteYN1 <- tbl_df(whiteYN) %>% 
  group_by(year, InGroup) %>% 
  summarise(howmany=sum(n)) %>% 
  mutate(frac=prop.table(howmany))

whiteplot <- ggplot(aes(x=year, y=howmany), data=whiteYN1) +
  geom_bar(aes(fill=InGroup), position="dodge", stat="identity") + 
  xlab("Year") + 
  ylab("Sample size") + 
  guides(fill=guide_legend(title=NULL)) + 
  guides(fill=guide_legend(title=NULL)) + 
  scale_fill_manual(values=colors2, labels=c("White", "Non-white"))
whiteplot
@

@@

<<OverallSeniors>>=
seniorGroup10 <- sotc10 %>% 
  filter(is.na(seniors) == FALSE) %>% 
  group_by(citystate, age >= 62, seniors) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)

@@

seniorGroup09 <- sotc09 %>% 
  filter(is.na(age) == FALSE) %>% 
  group_by(citystate, age >= 62, seniors) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(seniorGroup09) <- n1

seniorGroup08 <- sotc08 %>%
  group_by(citystate, age >=62, seniors) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2008)
names(seniorGroup08) <- n1

@@

<<kidsbyYear>>=
kidGroup10 <- sotc10 %>% 
  filter(children == "Yes" | children == "No", is.na(families_kids) == FALSE) %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)
names(kidGroup10) <- n1

kidGroup09 <- sotc09 %>% 
  filter(children == "Yes" | children == "No", is.na(families_kids) == FALSE) %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(kidGroup09) <- n1

@@

kidGroup08 <- sotc08 %>% 
  filter(children == "Yes" | children == "No", is.na(families_kids) == FALSE & families_kids != "(DK)" & families_kids != "(Refused)") %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2008)
names(kidGroup08) <- n1

@@

<<kidsSS,out.width="0.99\\linewidth",fig.width = 8.5,fig.height=4.5, fig.cap="Sample sizes for plots about meta-knowledge regarding the community as a place for families with young children.">>=
kidSS <- group_by(kidYears, year, InGroup) %>% 
  summarise(number=sum(num)) %>% 
  mutate(percentage=100*number/sum(number))

levels(kidSS$InGroup) <- c("Don't know", "Refused", "Childless households", "Families with children")
kidsplot <- ggplot(aes(x=year, y=number), data=kidSS)+geom_bar(aes(fill=InGroup), position="dodge", stat="identity")+xlab("Year")+ylab("Sample size")+guides(fill=guide_legend(title=NULL))+scale_fill_manual(values=colors2)
kidsplot 
@
